\textls[-0]{\thetitle}

### 2023-02-07 pldi2023/review-reply-densed.md

We thank all the reviewers for their critical and encouraging feedback.
Below we first address the main concerns. We might paraphrase

* (Paragraph 3/4 from Reviewer D)**The presentation of the calculus seems to be geared so that it closely mirrors an Agda formalization, in lieu of a more conventional presentation (e.g. substitutions are part of the syntax, and the calculus uses De Bruijn indices). Presumably due to space concerns, the formalization also lacks sufficient examples, which makes it hard to understand how things fit together: As one example, the description of the TM/WSUP rule does not mention the role of the signature $\tau$ in the rule, leaving the reader to figure out its purpose. It would be helpful to provide an concrete example of how this rule is used. I appreciate the effort to ground things via Figure 8's mapping from a few definitions of the STLC family to the core calculus. Unfortunately, this listing is effectively a partial definition, in that it contain several intermediate terms ($l_x$) whose definitions are not given. At the end of the day, I did not emerge from this section with a good intuition of how 'linkages' work and how they are used. In a similar vein, the proof of the key metatheoretic properties assumes that the reader is familiar with similar proofs, stating that the key technical device (an interpretation into a metalanguage) is largely similar to e.g. Altenkirch and Kaposi, without clearly explaining the extension. More worrisome, the full proof is *only* given in a notation heavy, Agda-style format, which is both unchecked by a machine and hard for humans to read.  While the formalization tries to rebut these concerns by acknowledging that the presentation will be dense to those 'without prior exposure to MLTT', but it would be more accurate to stipulate 'non-experts in MLTT'.**

<!-- I am not sure if I want to reply this or how I would like to respond this. 
      The only part I want to reply is 
        "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format, which is both unchecked by a machine and hard for humans to read."
      But my reply will certainly upgrade to a notation war that leads to nowhere.

    The real reply I want to give is: For example, For normalization of MLTT, given Abel's thesis (written in conventional style) and Kaposi's thesis (written in fake Agda style), I will believe the latter proof even in pen-and-paper form. This is due to the personal taste and numerous error I have personally made when I use latex/human language to lay out any mathematical object. So I simply believe fake-Agda is better for Human development (at least for myself).

    Apart from the above sentence, I think I can make no reply but really just do some modification suggestion on the paper
 -->

* (Paragraph 6) Reviewer D gives an example that will lead to awkwardness in our paper, because it require `coe` or equality transpose between propositionally equal types (due to the loss of judgemental(definitional) equality).

<!-- He got a point. To avoid transpose we need overridablity/pins which is not properly covered in the paper. -->

* (Paragraph 9) Reviewer D suggests the inclusion of `fdestruct` tactic

<!-- Generally speaking this is either not possible or trivial.
      For example, for overridable/pins it is trivial,
                  for non-overdiable stuff, it is not possible because the inductive type is extensible
    But I don't know If I want to tell the reviewer at all  -->

* **What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching? Is it rejected? Can it be defined, but only works for STLC.tm? In the latter case, how does the plugin prevent other extensible definitions from referring to it?** (Review D)

* **"In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?** (Review D)

```C
// For example
Family A {Field a; Field b; Field c;}

Family B extends A {
  // Inherit a. // this command line will make a difference once uncommented. Because the fields are order sensitive. In the case when a2's definition is dependent on a, omitting the inherit won't type check a2 
  // Thus this ``annotation`` require the programmer's effort. 
  Field a2.
}
```

* (Reviewer C) **The utility of the FMLTT formalization in this paper is unclear to me. Don't you already have consistency and canonicity via the translation to Coq demonstrated by FPOP? Or put another way, instead of defining FMLTT directly, couldn't you translate it to MLTT following the approach in FPOP?**

The technical detail here is that, FMLTT also has this unconventional Wtype formulation, which is closer to the inductive type in the surface syntax for Coq and Agda.

Our translation doesn't translate the unconventional Wtype formulation back to the conventional Wtype one. So the core calculus after translation is MLTT + this unconventional Wtype.

We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text.

* **looking at figure 2 this extension involves adding cases to the term language and adjusting the proofs, but not adding e.g. new things to the typing judgment.  Imagine we extend STLC to System F, using two type contexts, one for term variables, the other for type variables. Can such an extension be defined?**

* (Reviewer B)  **Would/does your approach also work in the presence of other approaches to binding, e.g. de Bruijn or locally nameless?**

* (Reviewer B) **looking at figure 2 this extension involves adding cases to the term language and adjusting the proofs, but not adding e.g. new things to the typing judgment.  Imagine we extend STLC to System F, using two type contexts, one for term variables, the other for type variables. Can such an extension be defined?** (Reviewer B also asked a question on extending the feature of reference)

To extend typing judgement is actually adding new indices for an existent inductive type, which is not covered in our paper. This not only happens for polymorphism extension, but also for the case of extending with the heap -- since reduction relation will have a new information on heap.

This is a quite challenging problem we are facing right now as well. The main challenge here for us is how to achieve code/proof reuse after this novel extension.

* (Reviewer B) **the fake-Agda formalization of translation is at any rate not a direct proof of correctness of the plugin implementation.**

We consider the formalized translation as an oversimplified fundation of the plugin implementation, also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text.

* (Reviewer B) **Where does the A in rules like TyEq/PK/Add come from?**

At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
In practice, the plugin implementation always choose a "default" one. We didn't formalize the "default" one as it is usually just making inductive type into an abstracted form and make sure other parts stay the "same" (I.e. family {Nat : W(O, S); pred : Nat -> Nat} is abstracted into {Nat : U; O : Nat; S : Nat -> Nat; pred : Nat -> Nat}).

* (Reviewer A/B) **In the conclusion of rule Tyeq/Casety, R should be T?** 
Yes. Sorry for the confusing typo.

* (Reviewer A) **In terms of ergonomics, I wonder how powerful and usable is the FInduction construct is. In Coq proofs, it's not rare to use induction tactics inside a larger proof. Will users need to outline those into their explicit own lemmas?**

Currently at plugin level, we only support layout the subcases in a lemma forms. 
  I think the design challenge of making FInduction-alike structure inside proof script, is mainly how extension happens -- the plugin either needs to manipulate the proof term or manipulate the proof script itself. 
  A. The former idea needs either 
      (1) extend the coq's (proof assistant's) kernel (i.e. make the pattern matching facility extensible) or 
      (2) maniuplate the AST of the proof term
  B. the later idea

  B breaks the expectation of incremental compilation a bit. 
  A.(1) and A.(2) both require sophisticated engineering effort. A(1) will also risk sabtaging the kernel itself and expanding the trusted base.
  Modulo the interesting semantic on the pattern matching itself (i.e. if we always consider the pattern matching can be elaborated to the usage of eliminator), I believe the current prototype is exploring the same semantic. So we didn't choose to explore this direction but this is definitely more practical.

* (Reviewer A) **Why is W(t) a term and not a type? What are bold-W and bold-P needed for?**

bold-W is a typo. We forgot to delete it.

ℙ is a technical device of the system mentioned on Line 788. Its functionality is to transform a linkage(overridable/extensible family) into a module(sigma type). The reason we need it is because, proper abstraction `A` cannot happen on linkage (𝕃) but only on a module (ℙ). (i.e. contrary to Line 788, the `Γ, 𝕃(𝜎) ⊢ s : A[p¹]` is usually unprovable for nontrivial `A`)

<!-- Do I want to expand it? -->

### 2023-02-07 pldi2023/review-reply-densed.md

We thank all the reviewers for their critical and encouraging feedback.
Below we first address the main concerns, which might be paraphrased for the space.

* (Reviewer A) "In terms of ergonomics, I wonder how powerful and usable" this is ... ; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."

<!-- I am not sure if the following respond is appropriate -->

As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level. We only have `fdiscriminate` and `finjection` for intermediate fact. We acknowledge this limitation of our current research and are trying to advance more in this future direction.

We believe the design challenge of making FInduction-alike structure inside proof script, is mainly how extension happens -- the plugin either needs to manipulate the proof term or manipulate the proof script itself. 
A. The former idea needs either 
    (1) extend the Coq's kernel (i.e. make the Coq's pattern matching facility extensible) or 
    (2) maniuplate the AST of the proof term
B. the latter will manipulate the AST of the proof script

B will break incremental compilation "a bit more" (i.e. we cannot have incremental compilation for each recursor). We believe the promising future lies in the direction of A(1), by enhance proof assistants with native extensible pattern matching instead of this current encoding. As Reviewer D mentioned, there is much work to be done.

<!-- 
// I am not sure if I want to include the following

Reviewer D points out the current lack of "fdestruct". In fact, we believe "fdestruct" doesn't exist. 
Both "finjection" and "fdiscriminate" works on arbitrary extensible inductive type but "fdestruct" can only work on extensible inductive type with fixed number of constructors -- that is a contradiction because all the extensions right now are about extending with new constructors.

One way to have destruct tactic usable is to use overridable/pins. This is not completely covered in Section 3.3, but this is used when we want overridable term to be transparent and thus an inductive type and "destruct"-able

 -->

* "Seems related to type theories for modules, but doesn't discuss the connection"

Finally, I note that a family has similarities with an ML-style module,
including the fact that some of its members are opaque while others are
abstract. This is modelled through the use of singleton types, just as
dependent type theories for modules do (e.g., Dreyer, Crary, Harper, POPL
2003), which are also extensions of MLTT. Furthermore, there has been work on
extending ML modules with mixin composition (Dreyer, Rossberg, ICFP 2008).
ML-style modules are the basis of Coq's module system as well, so I would have
expected some comparison with this line of work.

  > We need to say something about "A type system for higher-order modules"

* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition

Thanks for this feedback! We will definitely reorganize the whole structure for Section 5 to achieve better acessibility.

### Review B

* (Reviewer B)"One thing I don't really get from the rules/discussion (including the
  appendix) is the exact mechanism for choosing the A in rules like TyEq/PK/Add
  and LSig/Add where A is supposed to abstract out the possibly-extensible types
  in deciding what to make available to later definitions. Is A always going to
  be uniquely determinable from s? (or if not, does the nondeterminism matter?)
  Is it up to the user (in this case, the plugin) to get this right, or do the
  rules somehow make sure that the right things are abstracted in a way I can't
  see?"

At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type opauqe (Line 791) and make sure other parts stay the "same" 
(I.e. s maps {Nat : W(O, S); pred : Nat -> Nat} into {Nat : U; O : Nat; S : Nat -> Nat; pred : Nat -> Nat}). It is possible to design an interface where the programmer can choose which `A` they would love to abstract to, but we haven't found any use case for this functionality yet.

* (Reviewer B/D) "I find it somewhat surprising that neither the FPOP plugin / implementation
  itself, nor the code of the case studies, is provided."

We will definitely make the FPOP plugin/implementation available.

* (Review B) "I think the paper would also be improved by delineating the limitations of the
  current implementation / metatheory.."; (Reviewer C) "Be upfront about this limitation instead of hiding it in the section on case studies."; (Reviewer D) "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."; (Reviewer D) "It would be nice to add an explicit comment about the trusted code bases of any developments"

Thanks for the feedback! We will definitely explicitize the limitation of the current implementation and meta-theory in the final revision. Especially the problem concretized and emphasized by Reviewer D involving equality coercion (the pervasive propositional equality on extensible inductive family lead to unnecessary equality coercion e.g. when `denoteTy t` is not definitionally equal to `nat`).

We believe this can lead to exciting future work on both theory and practice -- a normaliation algorithm for an extended calculus admitting more property on extensible inductive type; and a proper proof assistant acknowledging native extensible inductive type instead of our current encoding by the plugin.

And yes, in the final revision we will explicitize the trusted codebase as our plugin only translate the code into the vanilla Coq term without expanding the trusted base.

### Review C

* "The utility of the FMLTT formalization in this paper is unclear to me. Don't
  you already have consistency and canonicity via the translation to Coq
  demonstrated by FPOP? Or put another way, instead of defining FMLTT directly,
  couldn't you translate it to MLTT following the approach in FPOP?"

@@

The technical detail here is that, FMLTT also has this unconventional Wtype formulation, which is closer to the inductive type in the surface syntax for Coq and Agda.

Our translation doesn't translate this unconventional Wtype formulation back to the conventional one. So the core calculus after translation is MLTT + this unconventional Wtype. In that case, translation function as a guide to the plugin implementation, and pedantically is not enough to show the consistency/canonicitiy for FMLTT.

<!-- We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text. -->
### Review D
* (Reviewer D) 
* > This expert reviewer's main concerns are (1) the incredible terseness of the theory
  presentation and (2) a few current limitations (framed as future research challenges).
  I don't think there are misunderstandings, so the strategy is to acknowledge them
  and promise to somehow address them in the final revision.

<!-- I think 
  terseness is already covered above
  challenge:
  1. the equality coerciion thing, covered above
  2. the limitedness of the current tactics, covered in the first question

  // TBH I don't know what are other tactics people usually use can be added here... But it is future work. I think this question

  // so I think there is no big questions from Reviewer D anymore.

-->

* Reviewer D reviewer also has some minor questions. Optionally respond to them.

* Reviewer A asks five questions in the "comments to authors" section;
  optionally answer them towards the end of the response letter.

### 2023-02-07 pldi2023/review-reply-densed.md

We will definitely make the FPOP plugin/implementation available.

* (Review B) "I think the paper would also be improved by delineating the limitations of the
  current implementation / metatheory.."; (Reviewer C) "Be upfront about this limitation instead of hiding it in the section on case studies."; (Reviewer D) "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."; (Reviewer D) "It would be nice to add an explicit comment about the trusted code bases of any developments"

Thanks for the feedback! We will definitely explicitize the limitation of the current implementation and meta-theory in the final revision. Especially the problem concretized and emphasized by Reviewer D involving equality coercion (the pervasive propositional equality on extensible inductive family lead to unnecessary equality coercion e.g. when `denoteTy t` is not definitionally equal to `nat`).

We believe this can lead to exciting future work on both theory and practice -- a normaliation algorithm for an extended calculus admitting more convertibility on elimination of the extensible inductive type; and a proper proof assistant acknowledging native extensible inductive type (instead of our current encoding by the plugin), and thus support the reduction of open terms properly.

### 2023-02-07 pldi2023/review-reply-densed.md

We thank all the reviewers for their critical and encouraging feedback.
Below we first address the main concerns, which might be paraphrased for the space.

* (Reviewer A) "In terms of ergonomics, I wonder how powerful and usable" this is ... ; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."

<!-- I am not sure if the following respond is appropriate -->

<!-- 
we hypothesize that nested induction can be lifted to the top level and then it will generate extra proof obligation balabla.

requires significant engineering effort

@@

 -->

As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level. We only have `fdiscriminate` and `finjection` for intermediate fact. We acknowledge this limitation of our current research and are trying to advance more in this future direction.

<!-- 
Doesn't seem very relevant for our paper rebuttal

We believe the design challenge of making FInduction-alike structure inside proof script, is mainly how extension happens -- the plugin either needs to manipulate the proof term or manipulate the proof script itself. 
A. The former idea needs either 
    (1) extend the Coq's kernel (i.e. make the Coq's pattern matching facility extensible) or 
    (2) maniuplate the AST of the proof term -> this will need to be shown at proof-script level
B. the latter will manipulate the AST of the proof script

B will break incremental compilation "a bit more" (i.e. we cannot have incremental compilation for each recursor). We believe the promising future lies in the direction of A(1), by enhance proof assistants with native extensible pattern matching instead of this current encoding. As Reviewer D mentioned, there is much work to be done. -->

<!-- Basically the generated proof term, we look at each pattern matching point
      and trying to give further proof obligation
    // concretize this idea
 -->

<!--

// the reviewer is actually asking nested pattern matching stuff here
// I am not sure if I want to include the following

Reviewer D points out the current lack of "fdestruct". In fact, we believe "fdestruct" doesn't exist.

Both "finjection" and "fdiscriminate" works on arbitrary extensible inductive type but "fdestruct" can only work on extensible inductive type with fixed number of constructors -- that is a contradiction because all the extensions right now are about extending with new constructors.

@@

Finally, I note that a family has similarities with an ML-style module,
including the fact that some of its members are opaque while others are
abstract. This is modelled through the use of singleton types, just as
dependent type theories for modules do (e.g., Dreyer, Crary, Harper, POPL
2003), which are also extensions of MLTT. Furthermore, there has been work on
extending ML modules with mixin composition (Dreyer, Rossberg, ICFP 2008).
ML-style modules are the basis of Coq's module system as well, so I would have
expected some comparison with this line of work.

  > We need to say something about "A type system for higher-order modules"

* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition

<!-- I love you chatgpt -->

Thank you for your feedback. We will promptly reorganize the structure of Section 5 to improve accessibility
<!-- sympethsize, dense ,hard to read, not giving too much

    we will expand  it in the main text to clarify the question the reviewers have, and also use appendices to explain the  formalization  in greater details 
         -->

@@

### Review B

* (Reviewer B)"One thing I don't really get from the rules/discussion (including the
  appendix) is the exact mechanism for choosing the A in rules like TyEq/PK/Add
  and LSig/Add where A is supposed to abstract out the possibly-extensible types
  in deciding what to make available to later definitions. Is A always going to
  be uniquely determinable from s? (or if not, does the nondeterminism matter?)
  Is it up to the user (in this case, the plugin) to get this right, or do the
  rules somehow make sure that the right things are abstracted in a way I can't
  see?"

At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type opauqe (Line 791) and make sure other parts stay the "same".

<!-- Use example on Pg 16. Concrete run it -->
(I.e. s maps {Nat : W(O, S); pred : Nat -> Nat} into {Nat : U; O : Nat; S : Nat -> Nat; pred : Nat -> Nat}). It is possible to design an interface where the programmer can choose which `A` they would love to abstract to, but we haven't found any use case for this functionality yet.

@@

* (Reviewer B/D) "I find it somewhat surprising that neither the FPOP plugin / implementation
  itself, nor the code of the case studies, is provided."

We will certainly make the FPOP plugin/implementation available.

@@

* (Review B) "I think the paper would also be improved by delineating the limitations of the
  current implementation / metatheory.."; (Reviewer C) "Be upfront about this limitation instead of hiding it in the section on case studies."; (Reviewer D) "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."; (Reviewer D) "It would be nice to add an explicit comment about the trusted code bases of any developments"

Thank you for your feedback. We will clearly state the limitations of the current implementation and meta-theory in the final revision. In particular,

1. we will emphasize the issue raised by Reviewer D regarding equality coercion and the pervasive propositional equality on extensible inductive families. This can result in unnecessary equality coercion when `denoteTy t` is not definitionally equal to `nat`.

We believe that this opens up new opportunities for future work in both theory and practice,
(a). such as a normalization algorithm for an extended calculus that allows for more convertibility upon elimination of the extensible inductive type; 
(b). and a proper proof assistant that supports native extensible inductive types instead of the current encoding via our plugin, thus allowing for proper reduction of open terms. This will change the kernel of proof assistant for sure.

In the final revision, we will also clarify that our plugin only translates code into vanilla Coq terms without expanding the trusted codebase.

@@

### Review C

* "The utility of the FMLTT formalization in this paper is unclear to me. Don't
  you already have consistency and canonicity via the translation to Coq
  demonstrated by FPOP? Or put another way, instead of defining FMLTT directly,
  couldn't you translate it to MLTT following the approach in FPOP?"

The technical detail here is that, FMLTT also has this unconventional Wtype formulation 
<!-- explain how unconventional it is
      what conventional about unconventional Wtype
 -->
, which is closer to the inductive type in the surface syntax for Coq and Agda.

@@

Our translation doesn't translate this unconventional Wtype formulation back to the conventional one. So the core calculus after translation is MLTT + this unconventional Wtype. In that case, translation function as a guide to the plugin implementation, and pedantically is not enough to show the consistency/canonicitiy for FMLTT.

<!-- make it clear that we have MLTT + unconventional Wtype as target calculus -->

<!-- acknowledge the reviewer's  idea of translating unconventional Wtype into conventional Wtype -->

@@

-->

* Reviewer D reviewer also has some minor questions. Optionally respond to them.

* Reviewer A asks five questions in the "comments to authors" section;
  optionally answer them towards the end of the response letter.

* Reviewer B ask some questions as well

* (Reviewer B) contradictory feature like polymorphism + heap causing unsoundness
<!-- we can expect mixin will generate proof obligation that is not pluasible be proved

one fixpoint extension, the other termination, definitely will fail on the reducibility argument.
 -->

### 2023-02-07 pldi2023/review-reply-densed.md

We thank all the reviewers for their critical and encouraging feedback.
Below we first address the main concerns, which might be paraphrased for the space.

* (Reviewer A) "In terms of ergonomics, I wonder how powerful and usable" this is ... ; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."

As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level. We only have `fdiscriminate` and `finjection` for deducing intermediate fact. The key problem is the lack of certain nested pattern matching.

<!-- The idea is 
    fill in the new hole once pattern matching needs to extend the clauses
    lifting nested induction to top level is not necessary
 -->

We acknowledge this limitation of our current research and are trying to advance more in this direction. We hypothesize that code involving nested induction/pattern matching an be reused by fill in the hole after extension upon inductive type. The plugin will generate extra proof obligation to fill this hole. This certainly requires significant amount of engineering effort.

@@

Finally, I note that a family has similarities with an ML-style module,
including the fact that some of its members are opaque while others are
abstract. This is modelled through the use of singleton types, just as
dependent type theories for modules do (e.g., Dreyer, Crary, Harper, POPL
2003), which are also extensions of MLTT. Furthermore, there has been work on
extending ML modules with mixin composition (Dreyer, Rossberg, ICFP 2008).
ML-style modules are the basis of Coq's module system as well, so I would have
expected some comparison with this line of work.

  > We need to say something about "A type system for higher-order modules"

* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition

Thank you for your feedback. 
We sympathize with the general impression on Section 5, and we agree it is dense, hard to read and not giving too much inspiration for the general understanding of the whole paper.

We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization in greater details.

@@

<!-- We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text. -->
### Other questions

* Reviewer A asks five questions in the "comments to authors" section;
  optionally answer them towards the end of the response letter.

* Reviewer B ask some questions as well

* (Reviewer B) How will our plugin react when trying to mixin the contradictory features? For example, since STLC with polymorphism and  STLC + references both enjoying principal typing under HM type inference, but their composition doesn't.

We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.

Another exmaple would be extend *STLC and its termination proof* with general recursion feature. Our plugin will generate unprovable proof obligation for the reducibility argument.

### 2023-02-07 pldi2023/review-reply-densed.md

Finally, I note that a family has similarities with an ML-style module,
including the fact that some of its members are opaque while others are
abstract. This is modelled through the use of singleton types, just as
dependent type theories for modules do (e.g., Dreyer, Crary, Harper, POPL
2003), which are also extensions of MLTT. Furthermore, there has been work on
extending ML modules with mixin composition (Dreyer, Rossberg, ICFP 2008).
ML-style modules are the basis of Coq's module system as well, so I would have
expected some comparison with this line of work.

  > We need to say something about "A type system for higher-order modules"

* (Reviewer A/B/C/D) Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition

Thank you for your feedback. 
We sympathize with this impression on Section 5, and we agree it is dense, hard to read and not giving too much inspiration for the readers.

@@

* (Reviewer B/D) "I find it somewhat surprising that neither the FPOP plugin / implementation
  itself, nor the code of the case studies, is provided."

We will certainly make the FPOP plugin/implementation available.

* (Review B) "I think the paper would also be improved by delineating the limitations of the
  current implementation / metatheory.."; (Reviewer C) "Be upfront about this limitation instead of hiding it in the section on case studies."; (Reviewer D) "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."; (Reviewer D) "It would be nice to add an explicit comment about the trusted code bases of any developments"

Thank you for your feedback. We will clearly state the limitations of the current implementation and meta-theory in the final revision. In particular, we will emphasize the issue raised by Reviewer D regarding equality coercion and the pervasive propositional equality on extensible inductive families. This can result in unnecessary equality coercion when `denoteTy t` is not definitionally equal to `nat`.

@@

<!-- We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text. -->
### Other questions

* Reviewer A asks five questions in the "comments to authors" section;
  optionally answer them towards the end of the response letter.

* Reviewer B ask some questions as well

* (Reviewer B) How will our plugin react when trying to mixin the contradictory features? For example, since STLC with (DHM-)polymorphism and  STLC + references both enjoying type soundness, but their composition doesn't.

@@

We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.

Another exmaple would be extend *a family of STLC and its termination proof* with general recursion feature. Our plugin will generate unprovable proof obligation inside the reducibility argument for the fixpoint feature.

### 2023-02-07 pldi2023/review-reply-densed.md

We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization in greater details.

### Review B

* (Reviewer B)"One thing I don't really get from the rules/discussion (including the
  appendix) is the exact mechanism for choosing the A in rules like TyEq/PK/Add
  and LSig/Add where A is supposed to abstract out the possibly-extensible types
  in deciding what to make available to later definitions. Is A always going to
  be uniquely determinable from s? (or if not, does the nondeterminism matter?)
  Is it up to the user (in this case, the plugin) to get this right, or do the
  rules somehow make sure that the right things are abstracted in a way I can't
  see?"

At the metatheory level, this A is decided by the programmer and not uniquely determined as you observed. 
In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type into an opaque type (Line 791) and make sure other parts stay the "same".

Taking (Line 745) Figure. 8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A` has `tm : 𝕌` instead. `s` will make sure other fields stay the same type. To show this explicitly at the plugin level, we look at Figure 4 (Line 494) `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding to `tm : 𝕌`). With this interface `STLC°tm` we cannot pattern match any term of type `tm : 𝕌` thus doing abstraction successfully.

Generally speaking, all the (extensible) inductive type will be simply "wrapped" by a module type with its interface and only exposing the constructor (with no eliminators), just like how we generate `STLC°tm`.

@@

<!-- We consider the formalized translation as an oversimplified fundation of the plugin implementation(because the unconventional Wtype is closer to the surface syntax so we believe it is legit), also act as a guidance of the plugin implementation, and hope it function as part of the supplementary text for the reader confused about the description in the main text. -->
### Other questions

* Reviewer A asks five questions in the "comments to authors" section;
  optionally answer them towards the end of the response letter.

* Reviewer B ask some questions as well

* (Reviewer B) How will our plugin react when trying to mixin the contradictory features? For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.

### 2023-02-07 pldi2023/review-reply-densed.md

In the final revision, we will also clarify that our plugin only translates code into vanilla Coq terms without expanding the trusted codebase.

### Review C

* "The utility of the FMLTT formalization in this paper is unclear to me. Don't
  you already have consistency and canonicity via the translation to Coq
  demonstrated by FPOP? Or put another way, instead of defining FMLTT directly,
  couldn't you translate it to MLTT following the approach in FPOP?"

The technical detail here is that, FMLTT also has this unconventional Wtype formulation. While each classic Wtype is defined by a pair of type `A ⊢ B`, our unconventional Wtype is defined by a list of pair of types `Aᵢ ⊢ Bᵢ`, each pair corresponds to one constructor.

One way to interpret it is the classical formluation encoding this list using a big sum type. So the conventional and unconventional Wtype can be translated to each other.

However, our translation doesn't translate this unconventional Wtype formulation back to the conventional one. So the target calculus after translation is **MLTT + this unconventional Wtype**. In that case, translation can only function as a guide to the plugin implementation, and **pedantically** is not enough to show the consistency/canonicitiy for FMLTT because **pedantically** target calculus is not proven to be consistency/canonicity.

We agree with Reviewer B's insight -- we will get consistency/canonicitiy when we translate unconventional Wtype into the conventional one. The reason we didn't choose to do so because we expect this translation a lot more verbose than the current proof because of the simplicity of Wtype itself compared to the rich functionality provided by (fake-)Agda's Inductive Facility. We only use the latter when constructing consistency/canonicity model.

### Other questions

### 2023-02-07 pldi2023/review-reply-densed.md

However, our translation doesn't translate this unconventional Wtype formulation back to the conventional one. So the target calculus after translation is **MLTT + this unconventional Wtype**. In that case, translation can only function as a guide to the plugin implementation, and **pedantically** is not enough to show the consistency/canonicitiy for FMLTT because **pedantically** target calculus is not proven to be consistency/canonicity.

We agree with Reviewer B's insight -- we will get consistency/canonicitiy when we translate unconventional Wtype into the conventional one. The reason we didn't choose to do so because we expect this translation a lot more verbose than the current proof because of the simplicity of Wtype itself compared to the rich functionality provided by (fake-)Agda's Inductive Facility. We only use the latter when constructing consistency/canonicity model.

### Other questions

* (Reviewer A) Why is W(t) a term and not a type? What are bold-W and bold-P needed for?

In dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.

𝕎(t) is a typo and should be deleted. Thanks for pointing out.

ℙ is another technical detail of the system. Its functionality is to transform a linkage(overridable/extensible) into a module(sigma type). The reason it is irreplacable is because, proper abstraction cannot happen on linkage (𝕃) but only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ : A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial `A₆`.

@@

* (Reviewer B) How will our plugin react when trying to mixin the contradictory features? For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.

We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.

Another exmaple would be extend *a family of STLC and its termination proof* with general recursion feature. Our plugin will generate unprovable proof obligation inside the reducibility argument for the fixpoint feature.

* (Reviewer D) **What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching? Is it rejected? In the latter case, how does the plugin prevent other extensible definitions from referring to it?**

Coq will reject on the pattern matching on `tm`. For example, Figure 4, Line 494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the given inductive `tm`. The following field definition will be type-checked based on this interface, and for Coq it will only consider `tm` as an arbitrary type and will fail to pattern match it.

<!-- I am not sure if I want to mention overridable/pins  -->

### 2023-02-08 coqimpl.tex
fix typos and awkward wordings

Upon \lsti{End}~\lsti{subst}, a module type named \lsti{STLC\_subst} is generated.
%
As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
% \EDJ{If we don't allow further binding case handler(mentioned in your last reply), how do the user further bind the subst? I mean in the surface syntax}\YZreply{The position is that individual case handlers cannot be further bound, but the recursive function itself can be further bound by acquiring new case handlers.}
so its definition is not exposed to the fields that come after it.
Accordingly, the translation \lsti{STLC\_subst} merely declares the types of \lsti{subst}
and the equalities about its computational behaviors, leaving \lsti{subst} undefined
and the equalities unproven.
%
The equalities are stated in terms of the case handlers,
whose definitions \emph{are} available through the \lsti{self} parameter.
So Coq can simplify, for example, the type of \lsti{subst_tm_unit_eq} to
\lsti[basicstyle=\fontsize{8.5}{9},basewidth=.95ex]{\forall x t, self.subst self.tm_unit x t = self.tm_unit}.
These equalities about the computational bahaviors of \lsti{subst}
will be included and available for use in the translations of the
subsequent fields through their \lsti{self} parameters.

>>>>

Upon \lsti{End}~\lsti{subst}, a module type named \lsti{STLC\_subst} is generated.
%
As discussed in \cref{sec:latebind}, \lsti{subst} can be further bound,
% \EDJ{If we don't allow further binding case handler(mentioned in your last reply), how do the user further bind the subst? I mean in the surface syntax}\YZreply{The position is that individual case handlers cannot be further bound, but the recursive function itself can be further bound by acquiring new case handlers.}
so its definition is not exposed to the fields that come after it.
Accordingly, the translation \lsti{STLC\_subst} merely declares the types of \lsti{subst}
and the equalities about its computational behaviors, leaving \lsti{subst} undefined
and the equalities unproven.
%
The equalities are stated in terms of the case handlers,
whose definitions \emph{are} available through the \lsti{self} parameter.
So Coq can simplify, for example, the type of \lsti{subst_tm_unit_eq} to
\lsti[basicstyle=\fontsize{8.5}{9},basewidth=.95ex]{\forall x t, self.subst self.tm_unit x t = self.tm_unit}.
These equalities about the computational behaviors of \lsti{subst}
will be included and available for use in the translations of the
subsequent fields through their \lsti{self} parameters.
